---
title: "STAC67 Case Study: A model for predicting housing values in Boston"
date: "March 20 2019"
author: 
- Revon Villava-Rayen (1002987501)
- Sizhe Gao ()
- Alexander Nguyen ()
- Gopi Santhiran ()
- Yu Heng Wu (1003475330)
- Group 13 
output: 
  word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
#install.packages("olsrr", "AICmodavg")
require("mosaic")
require("tidyverse")
require("olsrr")
require("car")
require("AICmodavg")
```

```{r, echo=FALSE}
housing = read.csv(file="housing.proper.csv",header=F)
colnames(housing) <- c('CRIME', 'ZL', 'NR_PROP', 'CHR_V', 'NOX', 'ROOM', 'AGE', 'DIS', 'HWY', 'TAX', 'PT_RATIO', 'B', 'L_PER', 'MEDV')
```

#Abstract
The housing market of the city of Boston is a volatile market, as with any world reknowned city. There are many attempts to create a formula to predict housing prices; using predictor variables such as per capita crime rate by town, or average number of rooms per dwelling. Using R statistical analysis, this study aims to examine and determine which model best represents the correlation between housing prices and these influential factors.



#Background and Significance
Boston is the captial and largest city in Massachusetts. Founded in 1630 Boston is one of the oldest cities in America and the city played an important role in American History. As Boston is a world reknowned city, the housing market in Boston is expensive, with finding and owning property is an important financial goal for many that live within or near Boston. As Housing is an expensive financial venture it is optimal to buy houses at a low price point to maximize 

The model introduced in this report is used to predict the housing value in the Boston with the given predictor values 

The goal of this study is to obtain an appropriate model 


#Explanatory Data Analysis
This Data set contains 506 observations on 13 predictors variables, which are per capita crime rate by town, proportion of residential land zoned for lots over 25,000 sq. ft., proportion of non-retail business acres per town
Charles River dummy variable, nitric oxide concentration (parts per 10 million), average number of rooms per dwelling
proportion of owner occupied units built prior to 1940, weighted distances to five Boston employment centres
index of accessibility to radial highways, full-value property-tax rate per 10,000, pupil-teacher ratio by town, 1000(B − 0.63)^2 where B is the proportion of African Americans by town, a numeric vector of percentage values of lower status population.

Because of the rather long names, we will assign shorter names in capitals listed below: 

###CRIME - per capita crime rate by town
A crime rate describes the number of crimes per 100,000 of the population. The mean and mode of the crime rate is _____ and ______ respectively 

###ZL - proportion of residential land zoned for lots over 25,000 sq. ft.
Placeholder

###NR_PROP - proportion of non-retail business acres per town
Placeholder

###CHR_V - Charles River dummy variable
Placeholder

###NOX - nitric oxide concentration (parts per 10 million)
Placeholder

###ROOM - average number of rooms per dwelling
Placeholder

###AGE - proportion of owner occupied units built prior to 1940
Placeholder

###DIS - weighted distances to five Boston employment centres
Placeholder

###HWY - index of accessibility to radial highways
Placeholder

###TAX - full-value property-tax rate per 10,000
Placeholder

###PT_RATIO - pupil-teacher ratio by town
Placeholder

###B - 1000(B − 0.63)^2 where B is the proportion of African Americans by town
Placeholder

###L_PER - a numeric vector of percentage values of lower status population
Placeholder

```{r, echo=FALSE}
lm.r=lm(MEDV ~ CRIME + ZL + NR_PROP + CHR_V + NOX + ROOM + AGE + DIS + HWY + TAX + PT_RATIO + B + L_PER, data=housing)
mean(lm.r)
```



#Model 




