---
title: "STAC67 Case Study: A model for predicting housing values in Boston"
date: "March 20 2019"
author: 
- Revon Villava-Rayen (1002987501)
- Sizhe Gao ()
- Alexander Nguyen ()
- Gopi Santhiran (1001371534)
- Yu Heng Wu (1003475330)
- Group 13 
output: 
  word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
#install.packages("olsrr", "AICmodavg","MASS")
require("mosaic")
require("tidyverse")
require("olsrr")
require("car")
require("AICmodavg")
require("MASS")
```

```{r, echo=FALSE}
housing = read.csv(file="housing.proper.csv",header=T,",")
colnames(housing) <- c('CRIME', 'ZL', 'NR_PROP', 'CHR_V', 'NOX', 'ROOM', 'AGE', 'DIS', 'HWY', 'TAX', 'PT_RATIO', 'B', 'L_PER', 'MEDV')
```

#Abstract
The housing market of the city of Boston is a volatile market, as with any world reknowned city. There are many attempts to create a formula to predict housing prices; using predictor variables such as per capita crime rate by town, or average number of rooms per dwelling. Using R statistical analysis, this study aims to examine and determine which model best represents the correlation between housing prices and these influential factors.



#Background and Significance
Boston is the captial and largest city in Massachusetts. Founded in 1630 Boston is one of the oldest cities in America and the city played an important role in American History. As Boston is a world reknowned city, the housing market in Boston is expensive, with finding and owning property is an important financial goal for many that live within or near Boston.

As Boston currently faces a "growing housing challenge ... likely because younger residents are marrying later ... many of them are overloaded with college debt" (Bluestone). It is clear that buying a House is an expensive financial burden that less and less people are able to afford. Especially when a house must be able to fit the needs of the buyer as a living space. 

We propose this model with the given predictor variables to accurately predict housing values within Boston. By being able to purchase a house at an optimal time, it minimizes the financial burden of the purchase and also allows you to maximize any capital gains that may result from the changing value of the purchased house. Through this process, it is possible to maximize the value of your purchase and minimize the cost of buying a house which leads to more people being able to afford housing in the city of Boston. 

#Exploratory Data Analysis
This Data set contains 506 observations on 13 predictors variables, which are per capita crime rate by town, proportion of residential land zoned for lots over 25,000 sq. ft., proportion of non-retail business acres per town
Charles River dummy variable, nitric oxide concentration (parts per 10 million), average number of rooms per dwelling
proportion of owner occupied units built prior to 1940, weighted distances to five Boston employment centres
index of accessibility to radial highways, full-value property-tax rate per 10,000, pupil-teacher ratio by town, 1000(B − 0.63)^2 where B is the proportion of African Americans by town, a numeric vector of percentage values of lower status population.

Because of the rather long names, we will assign shorter names in capitals listed below: 

```{r, echo=FALSE, include = FALSE}
coloumnMeans <- colMeans(housing,na.rm = TRUE)
coloumnMeans

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
Mode(housing$CRIME)
Mode(housing$ZL)
Mode(housing$NR_PROP)
Mode(housing$CHR_V)
Mode(housing$NOX)
Mode(housing$ROOM)
Mode(housing$AGE)
Mode(housing$DIS)
Mode(housing$HWY)
Mode(housing$TAX)
Mode(housing$PT_RATIO)
Mode(housing$B)
Mode(housing$L_PER)
``` 
###CRIME - per capita crime rate by town
The crime rate, measured by number of crimes per 100,000 of the population of the suburb observed. The mean and mode of this crime rate is 3.613 and 0.01501 respectively 

###ZL - proportion of residential land zoned for lots over 25,000 sq. ft.
This is the measurement of land set aside for residential buildings that is over 25,000 sq ft. The mean and mode of this is 11.363 and 0 and respectively. 

###NR_PROP - proportion of non-retail business acres per town
This is the measurement of the amount of non-retail land in the town. The mean and mode is 11.136 and 18.1 respectively.

###CHR_V - Charles River dummy variable
The Charles River is a 129 km long river in eastern Massachueetts. This is a binary variable and 6.9% of houses live near this river.

###NOX - nitric oxide concentration (parts per 10 million)
Refers to the concentration in parts per 10 million of nitric oxide. Short-term exposure to nitric oxide can inflict irritation to the respiratory system, and long term exposure can cause asthma and other respiratory infections (National Library of Medicine, 2017), suggesting that a large concentration of nitric oxide can adversely affect housing value. The mean and mode of nitric oxide concentration are 0.554 and 0.538 respectively.

###ROOM - average number of rooms per dwelling
This predictor measures the average room per dwelling in the observed suburb. The mean and mode of the number of rooms is 6.284 and 5.713 respectively.

###AGE - proportion of owner occupied units built prior to 1940
The mean and mode of the proportion of owner occupied units built prior to 1940 is 68.574 and 100 respectively.

###DIS - weighted distances to five Boston employment centres
The mean and mode weighted distances to five Boston employment centres is 3.795 and 3.4952 respectively.

###HWY - index of accessibility to radial highways
Radial highways are high-capacity urban roads leading to or from an urban center. The mean and mode of its index of accessibility is 9.549 and 24.

###TAX - full-value property-tax rate per 10,000
The full-value property-tax rate, measured per 10,000, meaning the amount of tax paid on a cap of $10000. The mean and mode of this tax rate is 408.237 and 666 respectively.

###PT_RATIO - pupil-teacher ratio by town
This ratio is the total number of pupils enrolled at a level of education divided by the number of teachers at that same level. "It is generally assumed that a low pupil-teacher ratio signifies smaller classes, ... teacher to pay more attention to individual student"(UNESCO, 2018) and vice versa for a high pupil-teacher ratio. The mean and mode is 18.455 and 20.2 respectively.

###B - 1000(B - 0.63)^2 where B is the proportion of African Americans by town
1000(B - 0.63)^2 is the scaling of the proportion of African Americans by town to match the scale of the other measurements. B represents the proportion of African American living in the town. The mean and mode of the proportion of African Americans by town is 356.674 and 396.9 respectively.

###L_PER - a numeric vector of percentage values of lower status population
This predictor is the percentage of lower status population in each observation. The mean and mode is 12.653 and 8.05.




#Model 

#Co-Lineartiy 
```{r, echo=FALSE}
res = cor(housing)
round(res,3)
```


The R code shown below displays the step forward and backwards process as well as the final model we will be using. 
```{r, echo=FALSE}
#Take the first 354 observations for our model  
mod_housing <- housing[1:354,c(1:14)]
#fit the model with the best observations 
fit1 <- lm(MEDV ~ CRIME + CHR_V + ROOM + AGE + DIS + PT_RATIO + B, data=mod_housing)
summary(fit1) 

#using the StepAIC forward and bac kwards to find the best model
fit_select = lm(MEDV ~ CRIME + ZL + NR_PROP + CHR_V + NOX + ROOM + AGE + DIS + HWY + TAX + PT_RATIO + B + L_PER, data=mod_housing)
step = stepAIC(fit_select, direction = "both")
step$anova

#Using AIC to reduce the number of predictor variables 
chosen_model = lm(MEDV ~ CRIME + CHR_V + NOX + ROOM + AGE + DIS + TAX + PT_RATIO + B + L_PER, data=mod_housing)

#AIC for chosen model
SSRes_1 = sum(chosen_model$residuals^2)
n_1 <- 354
p_prime_1 <-length(chosen_model$coefficients)
AIC_1 <- n_1*log(SSRes_1) - n_1*log(n_1) + 2*p_prime_1
AIC_1
#Using AIC to reduce the number of predictor variables 

#AIC for reduced model
reduced_model = lm(MEDV ~ CRIME + CHR_V + ROOM + AGE + DIS + PT_RATIO + B, data=mod_housing)
SSRes = sum(reduced_model$residuals^2)
n <- 354
p_prime <-length(reduced_model$coefficients)
AIC <- n*log(SSRes) - n*log(n) + 2*p_prime
AIC


# Studentized deleted residuals
"So from our outlier Test with t-crit value of 3.925756, we see that only outlying data point is 162."
outlierTest(fit_select)

t <- rstudent(fit_select)
alpha <- 0.05
# Transpose the MEDV from a column vector into a row vector in order for length to work
myVar <- housing["MEDV"]
n <- length(t(myVar))
p_prime = length(coef(fit_select)) 
t_crit <- qt(1-alpha/(2*n),n-p_prime-1)
round(t,2)
"Below is the T crit value: "
t_crit
which(abs(t) > t_crit)
  
#model validation
"Model Validation"
validate_housing <-housing[,c("CRIME", "CHR_V", "ROOM" , "AGE" , "DIS", "PT_RATIO","B")]


Y_pred <- predict(fit1, validate_housing)
Y_obs <- housing["MEDV"]
n_star <- nrow(housing)
MSPE <- sum((Y_obs-Y_pred)^2/n_star)
MS_res <-(summary(fit1)$sigma^2)
MS_res
MSPE

#ploting out the table 
#plot(fit1$residuals~fit1$fitted.values, ylab = "Residuals", xlab = "Fitted Values", main = "Graph Title")
#abline(0,0)
  
#qqPlot(fit1)

#model validation
#validate_housing <-housing[356:506,c(1:14)]
#Y_pred <- predict(fit1,validate_housing)
#Y_obs <- housing
#n_star <- nrow(housing)
#MSPE <- sum((Y_obs-Y_pred)^2/n_star)
#MS_res <-(summary(fit1)$sigma^2)
#MS_res
#MSPE

#outliers 
#outlierTest(fit_select)

#Outliers for Y
#s <- summary(fit_select)
#n <- nrow(surgic_mod)

#round(rbind(select_criteria(lm(MEDV ~ CRIME + ZL + NR_PROP + CHR_V + NOX + ROOM + AGE + DIS + HWY + TAX + PT_RATIO + B + L_PER, #data=mod_housing), n, s))


#t <- rstudent(fit_select)
#alpha <- 0.05
#n <- nrow(MEDV)
#p_prime = length(coef(fit_select)) 
#t_crit <- qt(1-alpha/(2*n),n-p_prime-1)
#round(t,2)
#t_crit
#which(abs(t) > t_crit)

#Leverage Points


#Influential Observations

```

#Citations 
Bluestone, B. & Huessy, J. (2017). The Greater Boston Housing Report Card 2017 Ideas from the Urban Core - Responsive Development - as a Model for Regional Growth, 68


